The effect of linking assumptions and number of response options on inferred scalar implicature rate

### 2018-04-30 writeup.Rmd
added bda to discussion/analysis

```{r load_packages, include = FALSE}
library(grid)
library(gtable)
library(papaja)
library(tidyverse)
library(magrittr)
library(readr)
library(png)
library(jpeg)
library(lme4)
library(ggthemes)
library(forcats)
# library(brms)
theme_set(theme_bw(18))
```

@@

On the Traditional story, we are forced to say that a) not all participants calculated the implicature; or that b) some participants who calculated the implicature did not choose the anticipated response (i.e. “Wrong”)  due to some other cognitive reflex which ‘overrode’ the implicature; or some mixture of (b) and (c). We might similarly posit that one or both of these factors underlie the variation in the ternary, quatenary, and quinary conditions (e.g. why were participants roughly split between “Right” and “Kind of right” when the utterance was cat or dog and the card had a cat and a dog?). There is an established cottage industry of pragmatics research which approaches the problem of behavioral variation in exactly these ways, exploring the factors which “license” implicature calculation or “suppress” its anticipated behavioral expression (c.f. __), but the best we can hope for on this approach is an analysis which traces the general qualitative patterns in the data (and we should stress that even in this respect, the industry has yet to fully deliver the goods!).

@@

R(u, w, $\theta$) 
    = “Right” iff $S_1$(u | w) $>$ $\theta$
    = “Wrong” otherwise

@@

R(u, w, $\theta_1$ , $\theta_2$)  
    = “Right” iff $S_1$(u | w) $>$ $\theta_1$
    = “Neither” iff $\theta_1$ $>$ $S_1$(u | w) $>$ $\theta_2$  
    = “Wrong” otherwise

@@

R(u, w, $\theta_1$ , $\theta_2$, $\theta_3$)  
    = “Right” iff $S_1$(u | w) $>$ $\theta_1$
    = “Kinda Right” iff $\theta_1$ $>$ $S_1$(u | w) $>$ $\theta_2$  
    = “Kinda Wrong” iff $\theta_2$ $>$ $S_1$(u | w) $>$ $\theta_3$ 
    = “Wrong” otherwise

@@

R(u, w, $\theta_1$ , $\theta_2$, $\theta_3$. $\theta_4$)  
    = “Right” iff $S_1$(u | w) $>$ $\theta_1$
    =“Kinda Right” iff $\theta_1$ $>$ $S_1$(u | w) $>$ $\theta_2$ 
    = “Neither” iff $\theta_2$ $>$ $S_1$(u | w) $>$ $\theta_3$  
    = “Kinda Wrong” iff $\theta_3$ $>$ $S_1$(u | w) $>$ $\theta_4$ 
    = “Wrong” otherwise

\newpage

>>>>

```{r load_packages, include = FALSE}
library(grid)
library(gtable)
library(papaja)
library(tidyverse)
library(magrittr)
library(readr)
library(png)
library(jpeg)
library(lme4)
library(ggthemes)
library(forcats)
library(webppl)
# library(brms)
theme_set(theme_bw(18))
```

@@

On the Traditional story, we are forced to say that a) not all participants calculated the implicature; or that b) some participants who calculated the implicature did not choose the anticipated response (i.e. “Wrong”) due to some other cognitive reflex which ‘overrode’ the implicature; or some mixture of (b) and (c). We might similarly posit that one or both of these factors underlie the variation in the ternary, quatenary, and quinary conditions (e.g. why were participants roughly split between “Right” and “Kind of right” when the utterance was cat or dog and the card had a cat and a dog?). There is an established cottage industry of pragmatics research which approaches the problem of behavioral variation in exactly these ways, exploring the factors which “license” implicature calculation or “suppress” its anticipated behavioral expression (c.f. __), but the best we can hope for on this approach is an analysis which traces the general qualitative patterns in the data.

@@

R(u, w, $\theta$)

= “Right” iff $S_1$(u | w) $>$ $\theta$

= “Wrong” otherwise

@@

R(u, w, $\theta_1$ , $\theta_2$)

= “Right” iff $S_1$(u | w) $>$ $\theta_1$

= “Neither” iff $\theta_1$ $>$ $S_1$(u | w) $>$ $\theta_2$

= “Wrong” otherwise

@@

R(u, w, $\theta_1$ , $\theta_2$, $\theta_3$)

= “Right” iff $S_1$(u | w) $>$ $\theta_1$

= “Kinda Right” iff $\theta_1$ $>$ $S_1$(u | w) $>$ $\theta_2$

= “Kinda Wrong” iff $\theta_2$ $>$ $S_1$(u | w) $>$ $\theta_3$

= “Wrong” otherwise

@@

R(u, w, $\theta_1$ , $\theta_2$, $\theta_3$. $\theta_4$)

= “Right” iff $S_1$(u | w) $>$ $\theta_1$

=“Kinda Right” iff $\theta_1$ $>$ $S_1$(u | w) $>$ $\theta_2$

= “Neither” iff $\theta_2$ $>$ $S_1$(u | w) $>$ $\theta_3$

= “Kinda Wrong” iff $\theta_3$ $>$ $S_1$(u | w) $>$ $\theta_4$

= “Wrong” otherwise

```{r echo=TRUE, message=FALSE}
bda <- webppl(program_file = "models/webppl_models/bda_model.wppl")

binary_graph <- ggplot(filter(bda, Parameter == "binary_theta"), aes(x=value))+
  # facet_wrap(~Parameter, scales = "free") +
  geom_density() +
  scale_fill_grey(guide=FALSE) +
  theme(axis.text.x=element_text(angle=20,hjust=1,vjust=1)) +
  ggtitle("Threshold distributions, binary condition:")

ternary_graph <- ggplot(filter(bda, Parameter == c("ternary_theta1", "ternary_theta2")), aes(x=value))+
  facet_wrap(~Parameter, scales = "free") +
  geom_density() +
  scale_fill_grey(guide=FALSE) +
  theme(axis.text.x=element_text(angle=20,hjust=1,vjust=1)) +
  ggtitle("Threshold distributions, ternary condition:")

quatenary_graph <- ggplot(filter(bda, Parameter == c("quatenary_theta1", "quatenary_theta2", "quatenary_theta3")), aes(x=value))+
  facet_wrap(~Parameter, scales = "free") +
  geom_density() +
  scale_fill_grey(guide=FALSE) +
  theme(axis.text.x=element_text(angle=20,hjust=1,vjust=1)) +
  ggtitle("Threshold distributions, quatenary condition:")

quinary_graph <- ggplot(filter(bda, Parameter == c("quinary_theta1", "quinary_theta2", "quinary_theta3", "quinary_theta4")), aes(x=value))+
  facet_wrap(~Parameter, scales = "free") +
  geom_density() +
  scale_fill_grey(guide=FALSE) +
  theme(axis.text.x=element_text(angle=20,hjust=1,vjust=1)) +
  ggtitle("Threshold distributions, quinary condition:")
binary_graph
ternary_graph
quatenary_graph
quinary_graph
```
\newpage
